% #############################################################################
% Agradecimentos / Acknowledgments
% !TEX root = ../main.tex
% #############################################################################

First and foremost, words cannot express my gratitude enough to my thesis supervisor, Alberto Abad. For your support, both scientific and personal, for your trust, your passion and for being the best supervisor I could have imagined. Arriving in Lisbon might have initially felt like stepping into the unknown, but it turned out to be one of the most rewarding decisions I have ever made. In particular, thank you for creating a healthy and pleasant working and learning environment during the years of this thesis. 

Additionally, I extend my heartfelt gratitude to Mariana, Francisco, Catarina, Vânia, Carlos and João, my dear office-mates from office number 212. Even during the long-distance periods with our Zoom breakfasts, I learned a lot alongside all of you. I also would like to thank all the Speech and HLTmeet fellows, including Isabel, Rubén, Anna Maria, John, Gonçalo, and everyone in the HLT team. Thank you all for your patience and for taking the time to speak English when I am around. Obrigado todos! 

Thank you to all the members of my thesis jury for taking the time to carefully read my manuscript.

This thesis owes much to the TAPAS research project. Beyond being just a project, it introduced me to incredible individuals who have enriched my journey. I am profoundly grateful for the opportunity to be a part of this project, and I extend my heartfelt thanks to everyone involved. Thank you to Timothy, Enno, Sebastião, Zhengjun, Viviana, Julian, Tomás, Zhao, Camilo, Wei, Bence, Eugenia and all your respective supervisors. This adventure has filled me with enough ducks and memories to last a lifetime.

My thesis adventure began in Toulouse, particularly at Samova in IRIT. I would like to thank the people there for motivating me to apply for this thesis. Thanks to Thomas, Julie, Jérôme, Lucile, Julien, and everybody from ``La Salle Machine" during my master's internship. 

Of course, I am grateful to my friends, especially those who engaged in numerous discussions about my thesis, including Edouard and Hugo. I am also thankful to my university friends, Théo, Roy, Véronique, Damien, Sofiane and the Derick team. Despite missing some Frontonades, every reunion feels like I left just yesterday. Moreover, even though we're far apart in distance, I would like to have a special thank you for Corentin, for all the hours we've spent together online and in person. Each of these moments holds great value for me. Also thanks to Rudy for always providing new music to add to my playlist.

By moving to Lisbon, I also had the pleasure of making new friends along the way whom I would like to thank. A big thanks to Govianna, Sara, Fabian, Anne, Octavio, Valentina, Ricardo, Flo', Jean, Adri', Ting, Thelma and everyone else who crossed my way during these years of thesis work. During these years, I've had the privilege of meeting people who have truly helped me grow as a person. Merci Priska, for always being there for me and for attempting to help me to keep my plants alive. Kiitos Ella, for teaching me how to push myself out of my comfort zone and for trying to spot northern lights with me.  

I would like to express my gratitude to all the people who may not have been mentioned in these pages but have contributed, even from afar, to this thesis.

Obrigado, Patricia, for bringing laughter into my life every day and for sharing your love of Italian food with me. Writing this thesis would not have been possible without your constant support. Thank you for taking care of me and for cheering me on through the late nights and early mornings. Your presence has been a source of comfort, inspiration and motivation.

Finally, I also want to express my heartfelt gratitude to my family. Thank you to my parents, for your unwavering support and patience, and for allowing me to embark on this incredible journey to Portugal. To my little sister Emma, thank you for your immense talent and for enduring my presence all these years. To my twin sister Aude, thank you for the countless hours you've spent supporting me over the phone and for always handling logistics for me, especially when it comes to gifts. To Vimel, thank you for being an amazing brother-in-law and for taking such good care of Aude. To my aunts, Tato, Cookie and Tata Lolo, thank you for your love and support. To my cousins, Nicolas, Cécilia, and Camille, thank you for being a source of joy in my life. And to my grandparents, Mamie Bordeaux and Papy Daniel, thank you for your wisdom and love.

My sincere thanks to all of you!

\vspace{1cm}
Thomas Rolland
\vspace*{\fill}
\begin{flushright}
    À la mémoire de Papy René et Mamie Brest. Je vous aime et j'espère vous avoir rendu fier.
\end{flushright}